---
title: "52 Pickup"
author: "laz97"
date: '2022-03-02'
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

## Introduction

Given the Crypto-Twitter’s chatter about $52 being the new floor price for LUNA, this dashboard aims to establish if there is any basis for said claim.

In the spirit of reproducibility, a public API was used and code can be accessed here[].

The dashboard was built in R Markdown to take advantage of available packages in R, including ggplot2’s graphing capabilities.
Pricing data comes from GoinGecko’s public API, and was accessed using the geckor package in R.

## LUNA’s price over the past year

Before establishing a time period to study the claim that \$52 is the floor price of LUNA, we decided to start by looking at a full year of price history to determine when was the first (and last) time that LUNA’s price crossed the \$52 mark for a prolonged period of time. 

```{r setup ,results = 'hide', echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Libraries
library(tidyverse)
library(tidyquant)
library(geckor)
library(ggplot2)
library(plotly)
library(RColorBrewer)
library(ggdark)

#LUNA Price History
LUNA  <- coin_history(coin_id = c("terra-luna"),
                        vs_currency = "usd",
                        days = 365,
                        interval = NULL,
                        max_attempts = 3
) 

```



```{r fig.width=9, fig.height=6, fig.align='center', include = TRUE, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
#Plot LUNA
#Plot LUNA
plot1 <- LUNA %>%
  ggplot(., aes(x = timestamp, y = price, color = coin_id)) + 
  geom_line() +
  scale_color_brewer(palette = "Set3") +
  dark_theme_gray() +
  geom_hline(yintercept = 52, color = "#FFFFB3", linetype = "dashed", size = 0.5) +
  theme(legend.position = "none") +
   ggtitle("LUNA Price History - 1 Year") +
    xlab("Date") + 
    ylab("Price")

ggplotly(plot1)

```

As we can see in the above graph,  the first time that LUNA crossed the \$52 barrier was before mid-November, but it only stayed above it for a day. However, on December 1st it crossed the mark again and stayed above it for almost two months, marking the first sustained excursion to the above-52 territory. 
 
Given this, we chose to study LUNA's price since December first in an attempt to establish if there is any basis to the claim that \$52 is the new floor price for LUNA.


## LUNA’s price since December 1st

The following plot shows LUNA's price since December 1st, conveniently marking a horizontal line at the $52 level. 

```{r fig.width=9, fig.height=6, fig.align='center', include = TRUE, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
#LUNA since december 1st plot
plot2 <- LUNA %>%
  filter(timestamp >= as.Date("2021-12-01")) %>%
  ggplot(., aes(x = timestamp, y = price, color = coin_id)) + 
  geom_line() +
  scale_color_brewer(palette = "Set3") +
  dark_theme_gray() +
  geom_hline(yintercept = 52, color = "#FFFFB3", linetype = "dashed", size = 0.5) +
  theme(legend.position = "none") +
  ggtitle("LUNA Price History - Since 2021-12-01") +
  xlab("Date") + 
  ylab("Price")
  
ggplotly(plot2)
```


The table below shows the total number of trading days in the studied period, the days when the close was above $52, and the days when the close was below.


```{r fig.width=9, fig.height=6, fig.align='center', include = TRUE, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
#Summary
LUNA %>%
  filter(timestamp >= as.Date("2021-12-01")) %>%
  mutate(above = if_else(price >= 52, 1,0),
         below = if_else(price >= 52, 0,1)) %>%
  summarise(trading_days = n(), above_52 = sum(above), below_52 = sum(below)) %>%
  knitr::kable(.,col.names = c("#Trading days",
                               "#Days above 52",
                                "#Days below 52"))

```


As we can see, since LUNA first crossed the $52 mark for a sustained period of time, it has only gone below briefly, by (relatively) little and very few times.

This, without being strong evidence that $52 can act as a floor to LUNA's price going forward, at least provides a basis for the claim.