---
title: "ASA Dex Degens"
author: "laz97"
date: '2022-03-24'
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

## Intro
On this dashboard, we study how Algorand wallets behaved on DEXs.

First, we show the distribution of the number of swaps by wallet since the beginning of the year.

Then, we show the distribution of the number of assets that were used in swaps by  wallets in the same period.

We do so using Flipsidecripto’s algorand.swaps table taking into consideration only swaps with positive swap_from_amounts. The query used to extract the data can be found [here](https://app.flipsidecrypto.com/velocity/queries/9b4d3d24-10e9-4fec-94a2-1c9e4951b0fe)

The rest of the analysis was done in R and the code can be found [here]() for reproducibility purposes.


```{r setup ,results = 'hide', echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Libraries
library(pacman)
pacman::p_load(tidyverse, dbplyr, RSQLite, lubridate, openxlsx, here, jsonlite, forcats, scales, plotly, RColorBrewer, ggdark,  ggiraph)

#Custom function
GGplotStyleMapper <- function(DF, Variable_Column, Values, Styles, ID_Column = NA) {
  
  if(length(Values) > length(Styles)) {warning("More values than styles")}
  
  non_matching <- setdiff(Values, unique(DF[[Variable_Column]]))
  if(length(non_matching) > 0) {warning("The following specified values do not match values in DF column:", paste(non_matching, collapse = " "))}
  
  #Map styles
  mapped_styles_specified <- tibble(Value = Values, Style = Styles[1:length(Values)])
  
  mapped_styles_others <- tibble(Value = setdiff(unique(DF[[Variable_Column]]), Values),
                                 Style = rep(Styles[(length(Values)+1):length(Styles)], length.out = length(setdiff(unique(DF[[Variable_Column]]), Values))))
  
  output <- full_join(mapped_styles_specified, mapped_styles_others, by = c("Value", "Style")) %>%
    deframe(.)
  
  
  #Not tested
  if(!is.na(ID_Column)) {
    if(nrow(distinct(DF, c(ID_Column))) != nrow(distinct(DF, c(ID_Column, Variable_Column)))) {
      warning("ID column values do not match criteria of having one match each on the Variable column")
    }
    
    output <- distinct(DF, c(ID_Column, Variable_Column)) %>%
      select(ID = 1, Value = 2) %>%
      left_join(., output, by = "Value") %>%
      select(., - Value) %>%
      deframe(.)
  }
  
  output
  
}

#Import ASA DEX Degens
dex_degens <- jsonlite::fromJSON("https://api.flipsidecrypto.com/api/v2/queries/9b4d3d24-10e9-4fec-94a2-1c9e4951b0fe/data/latest") %>% 
  as_tibble(.) %>%
  select(SWAPPER, SWAPS, SWAPPED_FOR)

```

## Distribution of the number of swaps by wallet

Below, we show both the frequency distribution and an  “inverse” cumulative distribution of the number of swaps each wallet performed since the beginning of the year.


```{r fig.width=9, fig.height=6, fig.align='center', include = TRUE, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
#Frequency
p0 <- ggplot(dex_degens, aes(x=SWAPS)) + 
  geom_histogram_interactive(bins=18, aes(y = stat(count) / sum(count),tooltip = round(stat(count) / sum(count), digits = 2)), color = RColorBrewer::brewer.pal(3,"Set3")[[1]] ) +
  scale_x_continuous(trans = "log10", labels = label_comma()) +
  scale_color_brewer(palette = "Set3") +
  dark_theme_gray() +
  ggtitle("Frequency distribution wallets' swaps") +
  xlab("Wallets' swaps count") + 
  ylab("Frequency")
  
girafe(ggobj = p0)  
```
  
.

```{r fig.width=9, fig.height=6, fig.align='center', include = TRUE, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
#Cumulative
p1 <- dex_degens %>%
  mutate(PROB = cume_dist(SWAPS)) %>%
  arrange(SWAPS) %>%
  ggplot(., aes(x=PROB, y= SWAPS)) + 
  geom_step(color = RColorBrewer::brewer.pal(3,"Set3")[[1]]) +
  scale_y_log10(labels = label_comma()) +
  scale_color_brewer(palette = "Set3") +
  dark_theme_gray() +
  ggtitle("Inverse cumulative probability wallet's swap count") +
  xlab("Cumulative probability") + 
  ylab("Swaps count")

ggplotly(p1)


```

As we can see, up to 30% of the wallets only performed one swap in this period, while slightly more than half performed less than 3 swaps. 

Only about 20% of wallets performed more than 13 swaps, while the top 10% performed more than 35.

On the more extreme side of the distribution, the top 1% of wallets performed more than 334 swaps while the top 0.1% performed more than 2.226 swaps.

The top wallet made more than 300.000 swaps in this period.


## Distribution of the number of swapped-for assets by wallet

Below, we show both the frequency distribution and an  “inverse” cumulative distribution of the number of assets swapped for others that each wallet performed since the beginning of the year.


```{r fig.width=9, fig.height=6, fig.align='center', include = TRUE, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
#Assets swapped by wallet
#Frequency
p2 <- ggplot(dex_degens, aes(x=SWAPPED_FOR)) + 
  geom_histogram_interactive(aes(y = stat(count) / sum(count), tooltip = round(stat(count) / sum(count), digits = 2)), bins = 13, color = RColorBrewer::brewer.pal(3,"Set3")[[1]]) +
  scale_x_log10() +
  scale_color_brewer(palette = "Set3") +
  dark_theme_gray() +
  ggtitle("Frequency distribution wallets' swappped-for ASAs") +
  xlab("Wallets' swapped for ASAs") + 
  ylab("Frequency")
girafe(ggobj = p2) 
```
  
.
```{r fig.width=9, fig.height=6, fig.align='center', include = TRUE, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
#Cumulative
p3 <- dex_degens %>%
  mutate(PROB = cume_dist(SWAPPED_FOR)) %>%
  arrange(SWAPPED_FOR) %>%
  ggplot(., aes(x=PROB, y= SWAPPED_FOR)) + 
  geom_step(color = RColorBrewer::brewer.pal(3,"Set3")[[1]]) +
  scale_y_log10(labels = label_comma()) +
  scale_color_brewer(palette = "Set3") +
  dark_theme_gray() +
  ggtitle("Inverse cumulative probability wallet's swappped-for ASAs") +
  xlab("Cumulative probability") + 
  ylab("Swappped-for ASAs")

ggplotly(p3)
```

As we can see, more than  45% of the wallets only swapped 1 asset since the beginning of the period, while slightly more than two-thirds used only 2 assets. 

Only about 20% of wallets used more than 4 assets, while the top 10% swapped 9 or more.

On the more extreme side of the distribution, the top 1% of wallets swapped more than 49 assets while the top 0.1% swapped more than 160.

The top wallet swapped 363 ASAs in this period.

