---
title: "ASA Dex Degens"
author: "laz97"
date: '2022-03-24'
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

## Intro

In this dashboard, we study the most popular assets to swap to in Algorand’s various DEXs.

First, we show the 10 most popular assets and the total number of times they were swapped to since the beginning of the year on all DEXs.

Then, we segregate the most popular assets on each DEX and compare them.

We do so using Flipsidecripto’s algorand.swaps table taking into consideration only swaps with positive swap_from_amounts. The queries used to extract the data can be found [here](https://api.flipsidecrypto.com/api/v2/queries/b4bd85dd-ece8-4aac-aa21-7963e6f25ddf/data/latest) and [here](https://api.flipsidecrypto.com/api/v2/queries/94f7b61b-1c6c-4e54-ad13-e0c62241699a/data/latest).

The rest of the analysis was done in R and the code can be found [here]() for reproducibility purposes.


```{r setup ,results = 'hide', echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Libraries
library(pacman)
pacman::p_load(tidyverse, dbplyr, RSQLite, lubridate, openxlsx, here, jsonlite, forcats, scales, plotly, RColorBrewer, ggdark)

#Custom function
GGplotStyleMapper <- function(DF, Variable_Column, Values, Styles, ID_Column = NA) {
  
  if(length(Values) > length(Styles)) {warning("More values than styles")}
  
  non_matching <- setdiff(Values, unique(DF[[Variable_Column]]))
  if(length(non_matching) > 0) {warning("The following specified values do not match values in DF column:", paste(non_matching, collapse = " "))}
  
  #Map styles
  mapped_styles_specified <- tibble(Value = Values, Style = Styles[1:length(Values)])
  
  mapped_styles_others <- tibble(Value = setdiff(unique(DF[[Variable_Column]]), Values),
                                 Style = rep(Styles[(length(Values)+1):length(Styles)], length.out = length(setdiff(unique(DF[[Variable_Column]]), Values))))
  
  output <- full_join(mapped_styles_specified, mapped_styles_others, by = c("Value", "Style")) %>%
    deframe(.)
  
  
  #Not tested
  if(!is.na(ID_Column)) {
    if(nrow(distinct(DF, c(ID_Column))) != nrow(distinct(DF, c(ID_Column, Variable_Column)))) {
      warning("ID column values do not match criteria of having one match each on the Variable column")
    }
    
    output <- distinct(DF, c(ID_Column, Variable_Column)) %>%
      select(ID = 1, Value = 2) %>%
      left_join(., output, by = "Value") %>%
      select(., - Value) %>%
      deframe(.)
  }
  
  output
  
}

#Import assets to swap to
swap_to <- jsonlite::fromJSON("https://api.flipsidecrypto.com/api/v2/queries/b4bd85dd-ece8-4aac-aa21-7963e6f25ddf/data/latest") %>% 
  as_tibble(.) %>%
  select(SWAP_TO_ASSET_ID, ASSET_NAME, SWAPS, RANK)



#Assets to swap to by dex
swap_to_dex <- jsonlite::fromJSON("https://api.flipsidecrypto.com/api/v2/queries/94f7b61b-1c6c-4e54-ad13-e0c62241699a/data/latest") %>% 
  as_tibble(.) %>%
  select(SWAP_PROGRAM, SWAP_TO_ASSET_ID, ASSET_NAME, SWAPS, RANK)

```

## Top 10 assets to swap to

Below we show a table with the top 10 ASAs to swap to in Algorand since the start of the year and a chart with the swap count of each.

```{r fig.width=9, fig.height=6, fig.align='center', include = TRUE, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
#Top 10 assets to swap to table
swap_to %>%
  filter(RANK <= 10) %>%
  knitr::kable(.)

```

Unsurprisingly, the most popular asset is ALGO, followed not so closely by Yieldly and USDC. The next 7 assets had significantly fewer swaps and were swapped-to a similar number of times compared to each other.

## Top 10 assets to swap to by DEX

```{r fig.width=9, fig.height=6, fig.align='center', include = TRUE, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
#Top 10 assets to swap to chart
p0 <- swap_to %>%
  filter(RANK <= 10) %>%
  mutate(ASSET_NAME = fct_reorder(ASSET_NAME, SWAPS, .desc = TRUE)) %>%
  ggplot(aes(x= ASSET_NAME, y = SWAPS)) +
  geom_col(aes(fill = ASSET_NAME)) +
  scale_y_continuous(labels = label_comma()) +
  scale_color_brewer(palette = "Set3") +
  dark_theme_gray() +
  ggtitle("Top 10 swapped-to assets") +
  xlab("Asset") + 
  ylab("Swapped-to times")

ggplotly(p0)
```
  
## Top 10 assets to swap to by DEX

In this section, we compare the most popular ASAs to swap to in each different DEX.
 
Below, we show a table with the 10 most popular assets to swap to by DEX (ranked) and a chart of the number of times each ASA was swapped-to for each one.

```{r fig.width=9, fig.height=6, fig.align='center', include = TRUE, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
#Taable
swap_to_dex %>%
  filter(RANK <= 10) %>%
  select(RANK, SWAP_PROGRAM, ASSET_NAME) %>%
  pivot_wider(names_from = SWAP_PROGRAM, values_from = ASSET_NAME) %>%
  knitr::kable(., longtable = TRUE)

```



```{r fig.width=9, fig.height=9, fig.align='center', include = TRUE, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
#Chart function
ChartByDex <-  function(dex = "tinyman") {
  plot<- swap_to_dex %>%
  filter(RANK <= 10 & SWAP_PROGRAM == dex) %>%
  mutate(ASSET_NAME = fct_reorder(ASSET_NAME, SWAPS, .desc = TRUE)) %>%
  ggplot(aes(x= ASSET_NAME, y = SWAPS)) +
  geom_col(aes(fill = ASSET_NAME)) +
  scale_y_continuous(labels = label_comma()) +
  scale_color_brewer(palette = "Set3") +
  dark_theme_gray() +
  ggtitle(dex) +
  xlab("Asset") + 
  ylab("Swapped-to times")
  
  ggplotly(plot) %>%
    hide_legend() %>%
    add_annotations(
      text = dex,
      x = 0.5,
      y = 1,
      yref = "paper",
      xref = "paper",
      xanchor = "middle",
      yanchor = "top",
      showarrow = FALSE,
      font = list(size = 15)
    )}

dexs <- swap_to_dex %>%
  distinct(SWAP_PROGRAM) %>%
  pull()


top10_by_dex <- lapply(setNames(dexs, dexs), ChartByDex)

subplot(top10_by_dex[["algofi"]], top10_by_dex[["pactfi"]],top10_by_dex[["tinyman"]], top10_by_dex[["wagmiswap"]], nrows = 2) %>%
  layout(title = "Top 10 swapped-to assets - by DEX")

```

As we can observe, ALGO was the top ASA in three out of the four DEXs. Surprisingly, ALGO was not even among the top 10 in algofi. 

Another remarkable fact out of this table is that wagmiswap only has three possible assets to swap to (which checks out according to their website).

Finally, Yieldly, USDC and STBL were constants on all three other DEXs.

